---
title: "Assignment 3: Data Visualization (incarceration)"
author: "ShengyueGao"
date: "2/22/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Introduction
In this report, I will focus on how the population (and rate) in jail (and in prison) differ by races by making comparisons and showing trends. I will use the dataset named **incarceration_trends.csv**, which i sloaded in analysis.R file. For the "Summary" and "Trends over Time Chart" part, I will choose 6 columns that contain the data of _population in jail_ for 6 races, 6 columns that describe the _population in prison_, and 6 columns
that describe the _prison admission_ for different races. For the "Variable Comparison Chart," I will pick 6 columns that describe the _rate_ of population that go into jail for different races and compare them with each other. For the "Map" part, I will use the _total_jail_adm_dcrp_ column to show the jail admission in each state in the United States in 2018. The _year_, _state_, and _county_ column will also be used in some parts. The columns holding specific race can better correspond to the theme of this project, which is to see how data differ by race, so I chose these variables and measures.

### Summary
As shown in the analysis.R file, I calculated 5 values from the dataset. 1) The first one is highest_jail_pop, which is the highest total number of population in jail from 1970 to 2018. The White has the highest jail population, which is 8814789. 2) The second one is lowest_jail_pop, which is the lowest population in jail from 1970 to 2018. The Other race has the lowest jail population, which is 178066.7. 3) The third one is highest_prison_pop, which is the highest total number of population in prison from 1970 to 2018. The Black has the highest prison population, which is 12113359. 4) The fourth one is the lowest_prison_pop, which is the population in prison from 1970 to 2018. The Asian American / Pacific Islander has the lowest prison population, which is  110036. 5) The last one is highest_prison_adm_ave, which is the highest average number of population of prison admission each year. The Black has the highest number of jail admission each year in average, which is 121877.

### Trends over Time Chart

```{r, echo=FALSE}
source("../source/trends.R")
chart
```





